---
title: "Prep_problem_solving"
output: html_document
---

```{r}
library(tidyverse)
library(stringr)
library(devtools)
library(bibi2.0) 
library(magrittr)
library(ggplot2)
```
Import the CSV containing the taxa information. Then, string trim and set all text to lowercase. Next, create a unique_id field, which is a concatenation of event_id and sample_number.
```{r}
prep.df <- read.csv("ChessieBIBI_biology.csv")

prep.df <- prep.df %>%
  clean_data_frame() %>%
  mutate_all(. , .funs=tolower) %>%
  mutate(tsn_final = str_replace_all(tsn_final, "(?<![0-9])0+", "")) %>%
  unite(unique_id, event_id, sample_number, remove = FALSE)
```
Taxa roll-ups and exclusions. Also removing unique_ids with counts of < 70. Then, convert the Functional Feeding Group (FFG) and Habit codes to full names. This adds two new columns, *ffg* and *habit*, to **prep.df**, which will be used instead of *bibi_ffg* and *bibi_habit* in later scripts.
```{r}
prep.df <- prep.df %>%
  prep_taxa() %>%
  mutate(reporting_value = as.numeric(reporting_value)) %>%
  filter(sum(reporting_value) > 70)

prep.df <- prep.df %>%
  mutate(ffg = case_when(
    is.na(bibi_ffg) | bibi_ffg == "" ~ as.character(NA),
    bibi_ffg == "cg" ~ "gather",
    bibi_ffg == "pr" ~ "predator",
    bibi_ffg == "sc" ~ "scrape",
    bibi_ffg == "cf" ~ "filter",
    bibi_ffg == "sh" ~ "shred",
    bibi_ffg == "om" ~ "omnivore",
    bibi_ffg == "pa" ~ "parasite",
    bibi_ffg %in% c("pc", "pi") ~ "pierce",
    TRUE ~ "ERROR"),
  habit = case_when(
    is.na(bibi_habit) | bibi_habit == "" ~ as.character(NA),
    bibi_habit == "sp" ~ "sprawl",
    bibi_habit == "cn" ~ "cling",
    bibi_habit == "cb" ~ "climb",
    bibi_habit == "bu" ~ "burrow",
    bibi_habit == "sw" ~ "swim",
    bibi_habit == "sk" ~ "skate",
    TRUE ~ "ERROR"))
```
Perform general cleaning of the data frame, and use `select` to retain only the indicated columns. The resulting data frame is **taxa.df**, which will be the input to the rarefaction step.
```{r}
fill_taxa <- function(long.df, final.id.col, ...) {
  rank.quos <- rlang::quos(...)
  final.id.col <- rlang::enquo(final.id.col)

  rank.vec <- long.df %>%
    select(!!!rank.quos) %>%
    names()

  final.df <- long.df %>%
    tidyr::gather(rank, taxon, !!!rank.quos) %>%
    dplyr::mutate(rank = factor(rank, levels = rank.vec),
                  taxon = if_else(taxon == "", as.character(NA), taxon)) %>%
    dplyr::group_by(!!final.id.col) %>%
    tidyr::fill(taxon)  %>%
    dplyr::ungroup() %>%
    tidyr::spread(rank, taxon)

  return(final.df)
  }

taxa.df <- prep.df %>% 
  ungroup() %>% 
  select(unique_id, icprb_bioregion_id, station_id, sample_date, reporting_value, final_id, phylum, subphylum, class, subclass, 
         order, suborder, family, subfamily, tribe, genus, habit, ffg, bibi_tv, aspt) %>%   
  group_by_at(vars(-reporting_value)) %>%
  summarize(reporting_value = sum(reporting_value)) %>%
  ungroup() %>%
  fill_taxa(. , final_id, phylum:genus) %>%
  mutate(final_id = genus) %>% 
  group_by_at(vars(-reporting_value)) %>%
  summarize(reporting_value = sum(reporting_value)) %>%
  ungroup()
```
